Volunteers in the Clouds: A Web Based Architecture to Harness Volunteer Resources for Low Cost Distributed Evolutionary Computation

### 2015-12-16 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Reviewing English grammar from State of the Art. I left a couple of comments/questions

The web used as a resource for distributed computing has a
long history, and almost since the beginning it was used for
evolutionary computation even before JavaScript was
introduced as a browser-based scripting language in 1997. Before that
time other technologies,
 including Flash animations and VBScript, ActiveX or Java applets were
used; for instance, Java was pointed out in \cite{soares1998get} as a
``language for the
internet'', providing some advantages such as multi-architecture compatibility or
security mechanisms, and was proposed in \cite{chandy1996world} as the
basis for the creation of a world-wide peer to peer system using compiled Java
programs. The ATLAS system
\cite{Baldeschwieler:1996:TIG:504450.504482} also used Java together
with another language called Cilk; all these papers indicated that
the introduction of an architecture-independent system such as the
Java virtual machine opened the doors for ultrascale, Internet-wide,
metacomputers. This paper also establishes the desirable features of
any {\em global computing} infrastructure, that can also be applied to
distributed evolutionary computation experiments that rely on
volunteer computing: scalability, heterogeneity, fault tolerance,
adaptive parallelism, volunteer's safety, anonymity, hierarchy, ease of use, and
reasonable performance. But hierarchy, this is, the capability of
preferentially using resources in the same organization, is a feature
that could be either disregarded or included within adaptivity, which
leaves us with eight desirable properties. 
% (Paloma) Re-write. For example, "But is hierarchy, this is, the
% capability of preferentially using resources in the same
% organization, the one that could be either disregarded or included
% in adaptivity, which leaves us with eight desirable properties." 
% (Paloma) And I re-wrote what I proposed yesterday.
% thanks :-) -JJ

The fulfillment of these properties could be achieved by using any of the
languages and associated technologies, mentioned above, provided that the user had
installed the required virtual machine plugin first. 
% (Paloma) These should be separate sentences
However, embedding a
distributed computation client in the browser has the advantage of
providing a single and universal way of accessing resources via an
URI, something that is not guaranteed by using desktop applications. 
Baratloo et al. tried in 1996 to create a framework with
the aforementioned qualities called {\em Charlotte}
\cite{baratloo1996charlotte}. This framework, which implemented a
distributed shared memory, was inspired by a previous work
on parallel computing libraries PVM and MPI and solved most of the issues by
using the Java virtual machine embedded in the browser,
which provided
security to the user, uniformity of computing environment
independently of the operating system, and access to a common data store in the
server. Since the Java virtual machine is a sandbox, it is safe to the user
and as it runs compiled bytecode it also guarantees a certain level of
performance.
Other issues such as scheduling, which are related to adaptive 
parallelism, were solved on the server. This
paper was later extended and published in a journal in 1999
\cite{baratloo1999charlotte}, but it is interesting to note that,
even in the early stages of the web, the possibility of
% (Paloma) Re-write. For example, "but it is interesting to note that, even in the early stages of the web, the possibility of..."
browser-based distributed computing was already researched.

Other systems, like JET, proposed by Soares et
al. \cite{soares1998get}, also used Java. JET supports
the execution of parallel applications over the Internet and features
a comprehensive statistics collection subsystem, enabling its use for science, %(Mario) Was easier to change "allowing"
since it allows % (Paloma) Use synonyms, like "yield" or "permit" 
the collection of algorithm analytics by the
experiment designer, unlike the other systems mentioned
above. Other proposals contemporary to JET, such as {\em SuperWeb}
\cite{alexandrov1997superweb} are roughly a translation of classical
distribution techniques, with overly complicated architectures that include
brokers for resource discovery. SuperWeb is focused,
% (Paloma) Careful with starting using plural ("Other proposals", "are") and continuig with singular. If you are referring to "SuperWeb", change "It" for "SuperWeb"
however, in solving the
issues involved in browser-based computing and also in making a system
that is able to work across different types of computers, mentioning
SPARCstations and PCs with Windows-95. Interestingly enough, this
paper incorporates a discussion on the economic model, paving the way
for the for-profit distributed computing systems that arose later on. 
On the other hand, the potential for volunteer computing using
browsers was realized
later on \cite{sarmenta-bayanihan} as well as the potential of
user misconduct when having access to the source code.
% (Paloma) I think this sentence is extremely informal
\cite{sarmenta-sabotagetolerance}. However, these early efforts by
Sarmenta once again used Java and not JavaScript, making them % (Paloma) making them
less widespread.

@@

JavaScript can be used either for unwitting
\cite{unwitting-ec} or volunteer 
\cite{langdon:2005:metas,gecco07:workshop:dcor} distributed
evolutionary computation and it has been used ever since by several
authors, including more recent efforts
\cite{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210} 
that even
used the client's GPU \cite{duda2013gpu} or create an object-oriented
framework for evolution in the browser \cite{EvoStar2014:jsEO}. Many other researchers have
used Java \cite{chong:1999:jDGPi} and others have gone away from the
server-based paradigm to embrace peer to peer systems
\cite{jin2006constructing,10.1109/ICICSE.2008.99,DBLP:conf/3pgcic/GuervosMFEL12}. These computing
platforms avoid single points of failure (the server) but, since they
need a certain amount of infrastructure installed to start, the
threshold to join them is much lower.

@@

JavaScript is also compatible with the functional programming paradigm \cite{Cousineau1998,MacLennan1990,Thompson1996}.
According to \cite{swanresearch2015}, the benefit of using this kind of paradigm in
metaheuristics optimization is that it allows an improvement in communicability,
reproductibility, interoperability, automated assembly, knowledge engineering
and efficiency.

@@

On the other hand, we are also interested in measuring the performance
of volunteer computing systems, an area in which there have been
relatively few efforts.
There were some initial attempts to avoid the differences in performance
that could be obtained from volunteers  by making
the algorithm adaptive to the kind of resources allotted to it
\cite{milani2004online}, which is actually not such a big problem in
algorithms such as the EA that can easily be
parallelized via population splitting or farming out evaluations to all
the nodes available. Lately, several approaches have focused on the
fault-tolerance of volunteer algorithms
\cite{gonzalez2010characterizing} which it can, of course, be studied in
the more general context of distributed computing 
\cite{nogueras2015studying} or including it in a more general study of the
performance of the EA itself
\cite{DBLP:journals/gpem/LaredoBGVAGF14}.

On the other hand, measuring the performance of the resulting metacomputer
involves understanding the dynamics of this kind of systems. Initial
work was done for peer to peer systems by Stutzbach et
al. \cite{stutzbach2006understanding} and extended to volunteer
computing by Laredo et al. \cite{churn08,laredo2008rcp}. 
%But the raw material of           
%volunteer computing, % (Paloma) Whith regard to?  Maybe it was colloquial/analogy
But some of the essential metrics in volunteer computing like the
number of users or the time spent by every one in the
computation in browser-based volunteer computing experiments, have
only been studied in a limited way in 
\cite{DBLP:journals/gpem/LaredoBGVAGF14} on the basis of a single
run. Studies using volunteer computing platforms such as SETI@home
\cite{javadi2009mining} found out that the Weibull, log-normal and
Gamma distribution 
modeled quite well the availability of resources in several clusters
of that framework; the shape of those distributions is a skewed bell
with more resources in the {\em low} areas than in the high areas:
there are many users that give a small amount of cycles, while there
are just a few that give many cycles. This is in concordance with the
results obtained in \cite{agajaj}. Although SETI@home works
at a larger scale in both quantity of volunteers and the scope of projects, 
our initial tests indicate that
this kind of distribution might be the most appropriate to describe
user participation in several different metrics. However, in this
paper we will perform different experiments in different setups and
will try to come up with a more precise model. Next we describe the
framework itself.

>>>>

The web used as a resource for distributed computing has a
long history, and almost since the beginning it was used for
evolutionary computation even before JavaScript was
introduced as a browser-based scripting language in 1997. Before that
time other technologies,
 including Flash animations and VBScript, ActiveX or Java applets were
used; for instance, Java was pointed out in \cite{soares1998get} as a
``language for the
internet'', providing some advantages such as multi-architecture compatibility or
security mechanisms, and was proposed in \cite{chandy1996world} as the
basis for the creation of a world-wide peer to peer system using compiled Java
programs. The ATLAS system
\cite{Baldeschwieler:1996:TIG:504450.504482} also used Java together
with another language called Cilk; all these papers indicated that
the introduction of an architecture-independent system such as the
Java virtual machine opened the doors for ultrascale, Internet-wide,
metacomputers. This paper % (Paloma) No really need for that "also", plus it complicates reading
establishes the desirable features of
any {\em global computing} infrastructure, that can also be applied to
distributed evolutionary computation experiments that rely on
volunteer computing: scalability, heterogeneity, fault tolerance,
adaptive parallelism, volunteer's safety, anonymity, hierarchy, ease of use, and
reasonable performance. But hierarchy, this is, the capability of
preferentially using resources in the same organization, is a feature
that could be either disregarded or included within adaptivity, which
leaves us with eight desirable properties. 
% (Paloma) Re-write. For example, "But is hierarchy, this is, the
% capability of preferentially using resources in the same
% organization, the one that could be either disregarded or included
% in adaptivity, which leaves us with eight desirable properties." 
% (Paloma) And I re-wrote what I proposed yesterday.
% thanks :-) -JJ

The fulfillment of these properties could be achieved by using any of the
languages and associated technologies, mentioned above, provided that the user had
installed the required virtual machine plugin first. 
% (Paloma) These should be separate sentences
However, embedding a
distributed computation client in the browser has the advantage of
providing a single and universal way of accessing resources via an
URI, something that is not guaranteed by using desktop applications. 
Baratloo et al. tried in 1996 to create a framework with
the aforementioned qualities called {\em Charlotte}
\cite{baratloo1996charlotte}. This framework, which implemented a
distributed shared memory, was inspired by a previous work
on parallel computing libraries PVM and MPI and solved most of the issues by
using the Java virtual machine embedded in the browser,
which provided
security to the user, uniformity of computing environment
independently of the operating system, and access to a common data store in the
server. Additionally, since the Java virtual machine is a sandbox, it was safe to the user
and as it runs compiled bytecode it also guaranteed a certain level of
performance. % (Paloma) I've changed a bit this sentence to unify the use of past tense when talking about Charlotte
Other issues such as scheduling, which are related to adaptive 
parallelism, were solved on the server. This
paper was later extended and published in a journal in 1999
\cite{baratloo1999charlotte}, but it is interesting to note that,
even in the early stages of the web, the possibility of
% (Paloma) Re-write. For example, "but it is interesting to note that, even in the early stages of the web, the possibility of..."
browser-based distributed computing was already researched.

Other systems, like JET, proposed by Soares et
al. \cite{soares1998get}, also used Java. JET supports
the execution of parallel applications over the Internet and features
a comprehensive statistics collection subsystem, enabling its use for science, %(Mario) Was easier to change "allowing"
since it allows % (Paloma) Use synonyms, like "yield" or "permit" 
the collection of algorithm analytics by the
experiment designer, unlike the other systems mentioned
above. Other proposals contemporary to JET, such as {\em SuperWeb}
\cite{alexandrov1997superweb} are roughly a translation of classical
distribution techniques, with overly complicated architectures that include
brokers for resource discovery. SuperWeb is focused,
% (Paloma) Careful with starting using plural ("Other proposals", "are") and continuig with singular. If you are referring to "SuperWeb", change "It" for "SuperWeb"
however, in solving the
issues involved in browser-based computing and also in making a system
that is able to work across different types of computers, mentioning
SPARCstations and PCs with Windows-95. Interestingly enough, this
paper incorporates a discussion on the economic model, paving the way
for the for-profit distributed computing systems that arose later on. 
On the other hand, the potential for volunteer computing using
browsers was realized
later on \cite{sarmenta-bayanihan} as well as the potential of
user misconduct when having access to the source code \cite{sarmenta-sabotagetolerance}. 
However, these early efforts by
Sarmenta once again used Java and not JavaScript, making them % (Paloma) making them
less widespread.

@@

JavaScript can be used either for unwitting
\cite{unwitting-ec} or volunteer 
\cite{langdon:2005:metas,gecco07:workshop:dcor} distributed
evolutionary computation and it has been used ever since by several
authors, including more recent efforts
\cite{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210} 
that even
used the client's GPU \cite{duda2013gpu} or create an object-oriented
framework for evolution in the browser \cite{EvoStar2014:jsEO}. Many other researchers have
used Java \cite{chong:1999:jDGPi} and others have gone away from the
server-based paradigm to embrace peer to peer systems
\cite{jin2006constructing,10.1109/ICICSE.2008.99,DBLP:conf/3pgcic/GuervosMFEL12}. These computing
platforms avoid single points of failure (the server) but, since they
need a certain amount of infrastructure installed to start, the
threshold to join them is much lower. % (Paloma) Do yo mean... higher? A low threshold means that is easy to happen something.

@@

JavaScript is also compatible with the functional programming paradigm \cite{Cousineau1998,MacLennan1990,Thompson1996}.
According to \cite{swanresearch2015}, the benefit of using this kind of paradigm in
metaheuristics optimization is that it allows an improvement in communicability,
reproductibility, interoperability, automated assembly, knowledge engineering,
and efficiency.

@@

On the other hand, we are also interested in measuring the performance
of volunteer computing systems, an area in which there have been
relatively few efforts.
There were some initial attempts to avoid the differences in performance
that could be obtained from volunteers  by making
the algorithm adaptive to the kind of resources allotted to it
\cite{milani2004online}. This is actually not such a big problem in
algorithms such as the EA that can easily be
parallelized via population splitting or by farming out evaluations to all
the nodes available. Lately, several approaches have focused on the
fault-tolerance of volunteer algorithms
\cite{gonzalez2010characterizing} which it can, of course, be studied in
the more general context of distributed computing 
\cite{nogueras2015studying} or be included in a more general study of the
performance of the EA itself
\cite{DBLP:journals/gpem/LaredoBGVAGF14}.

On the other hand, measuring the performance of the resulting metacomputer
involves understanding the dynamics of this kind of systems. Initial
work was done for peer to peer systems by Stutzbach et
al. \cite{stutzbach2006understanding} and extended to volunteer
computing by Laredo et al. \cite{churn08,laredo2008rcp}. 
%But the raw material of           
%volunteer computing, % (Paloma) Whith regard to?  Maybe it was colloquial/analogy
But some of the essential metrics in volunteer computing like the
number of users or the time spent by every one in the
computation in browser-based volunteer computing experiments, have
only been studied in a limited way in 
\cite{DBLP:journals/gpem/LaredoBGVAGF14} on the basis of a single
run. Studies using volunteer computing platforms such as SETI@home
\cite{javadi2009mining} found out that the Weibull, log-normal, and
Gamma distribution 
modeled quite well the availability of resources in several clusters
of that framework; the shape of those distributions is a skewed bell
with more resources in the {\em low} areas than in the high areas:
there are many users that give a small amount of cycles, while there
are just a few that give many cycles. This is in concordance with the
results obtained in \cite{agajaj}. Although SETI@home works
at a larger scale in both quantity of volunteers and the scope of projects, 
our initial tests indicate that
this kind of distribution might be the most appropriate to describe
user participation in several different metrics. However, in this
paper we will perform different experiments in different setups and
will try to come up with a more precise model. Next we describe the
framework itself.
